---
title: "WON_analysiswith2022data_16April_excessbiomassversion"
format: html
editor: visual
---

## WON paper with 2022 data

In this version of the analysis, I focus on "excess biomass" instead of overweight or obesity.

## Load data and packages:

```{r}
library(readr)
library(dplyr)
library(sn)
library(e1071)
library(fanplot)
library(forcats)
library(ggplot2)
library(readxl)
library(tidyr)
```

```{r}
#DATA BY AGE-SEX GROUP FOR POP, HEIGHT, BMI IN CATEGORIES
df_bmipopheight_2022 <- read_csv("Output/df_bmipopheight_2022_18plus.csv")
#PARAMETER ESTIMATES OF SPLIT-NORMAL DISTRIBUTIONS FOR BMI (BAYESIAN POSTERIOR MEANS)
bayesmod_hns <- readRDS("~/OneDrive/Documents/UAH/Weightofnations/Bayesian/twohalfnorms_results_newmid_rules_modeshift_2022_31Oct_pivotprior_pluslimit_plusnormal_updatedpriors.RDS")

```

## Preparation of data:

```{r}
#MERGE THE DATASETS ABOVE:
DF_ADULT<-df_bmipopheight_2022 
DF_ADULT_noNA<-DF_ADULT %>% filter(!is.na(`prev_less18.5`))
bayesmod_hns2<- bayesmod_hns %>% rename(SEX=sex)
DF_ADULT_bayesmod1<-DF_ADULT_noNA %>% full_join(bayesmod_hns2, by=c("name", "heightagegroups"="agegroup","SEX"))
DF_ADULT_bayesmod1$total_population<-DF_ADULT_bayesmod1$TOTPOP
#CALCULATE MEAN BMI IN EACH GROUP:
DF_ADULT_bayesmod1$BMIMEAN_hns<-DF_ADULT_bayesmod1$pivot+(DF_ADULT_bayesmod1$sigma2-DF_ADULT_bayesmod1$sigma1)*sqrt(2/pi)
#SET ORDER OF REGION VBL SO THAT IT PRINTS IN AN EASY ORDER TO TRANSFER TO TABLES:
DF_ADULT_bayesmod2 <- DF_ADULT_bayesmod1 %>%
  mutate(region = factor(region,
    levels = c("ASIA",
               "EUROPE",
               "AFRICA",
               "LATIN AM. CARIBBEAN",
               "NORTHERN AM.",
               "OCEANIA")
  ))
```

## TABLES

### Table 1: Excluded countries

```{r}
Table1 <- read_csv("Output/Table1_18plus_2022.csv")

Table1 %>% arrange(desc(total))

Table1 %>% summarise(world_popexcluded=sum(total,na.rm=T))
```

## 

### Table 2: Adult population, body mass and biomass

#### Current

##### Biomass estimates for each age-sex-country group

```{r}
#BIOMASS ESTIMATE BY AGE-SEX-COUNTRY GROUP:
DF_ADULT_bayesmod2[,paste(22,"biotot",sep="")] <- DF_ADULT_bayesmod2$total_population*DF_ADULT_bayesmod2$BMIMEAN_hns*((DF_ADULT_bayesmod2$meanheight/100)^2 + (DF_ADULT_bayesmod2$sdheight/100)^2) #see formula under table 1 for E(W)

#FILTER FOR ONLY THOSE AGE-SEX-COUNTRY GROUPS THAT HAVE BIOMASS ESTIMATE:
DF_ADULT_bayesmod <- DF_ADULT_bayesmod2 %>% filter(!is.na(`22biotot`))

#TOTAL BIOMASS BY COUNTRY
bm_bycountry<-DF_ADULT_bayesmod %>% group_by(name) %>% summarise(biotot_country_mil=sum(`22biotot`,na.rm=T)/1000)
```

##### Total biomass, population and avg body mass by WHO region

```{r}
#BIOMASS, POPULATION AND AVERAGE BODY MASS BY WHO REGION 
bmpop_region<-DF_ADULT_bayesmod %>% group_by(region) %>% summarise(biotot_region_mil=sum(`22biotot`)/1000, totpop_region_mil=sum(total_population,na.rm=T)/1000)
bmpop_region$averagebodymass<-bmpop_region$biotot_region_mil/bmpop_region$totpop_region_mil
bmpop_region
```

##### Total biomass, population and avg body mass for current WORLD

```{r}
#BIOMASS, POPULATION AND AVERAGE BODY MASS FOR WORLD
bmpop_world<-DF_ADULT_bayesmod %>% summarise(biotot_world_mil=sum(`22biotot`,na.rm=T)/1000, totpop_world_mil=sum(total_population,na.rm=T)/1000)
bmpop_world$averagebodymass<-bmpop_world$biotot_world_mil/bmpop_world$totpop_world_mil
bmpop_world
```

#### Optimal scenario (& differences in table 2)

##### Apply BMI distribution of Japan

```{r}
#JAPAN SCENARIO...
Japan<-DF_ADULT_bayesmod %>% filter(name=="Japan") %>% select(c(heightagegroups,SEX,BMIMEAN_hns))
Japan_re<-Japan %>% rename(BMIMEAN_hns_JAPAN=BMIMEAN_hns)

DF_ADULT_bayesmod_japan<-DF_ADULT_bayesmod %>% left_join(Japan_re,by=c("heightagegroups","SEX"))
```

##### Biomass estimates for each age-sex-country group under S1

```{r}
#TOTAL BIOMASS BY EACH AGE-SEX-COUNTRY GROUP WITH BMI DIST OF JAPAN
DF_ADULT_bayesmod_japan[,paste(22,"biotot_JAPAN",sep="")] <- DF_ADULT_bayesmod_japan$total_population*DF_ADULT_bayesmod_japan$BMIMEAN_hns_JAPAN*((DF_ADULT_bayesmod_japan$meanheight/100)^2 + (DF_ADULT_bayesmod_japan$sdheight/100)^2) #see formula under table 1 for E(W)
```

##### Total biomass and avg body mass by WHO region under S1 - and difference from current world

```{r}
#BIOMASS AND AVERAGE BODY MASS BY WHO REGION UNDER S1 - AND DIFFERENCE FROM CURRENT
bmpop_region_japan<-DF_ADULT_bayesmod_japan %>% group_by(region) %>% summarise(biotot_region_japan_mil=sum(`22biotot_JAPAN`,na.rm=T)/1000, totpop_region_japan_mil=sum(total_population,na.rm=T)/1000)
bmpop_region_japan$averagebodymass_japan<-bmpop_region_japan$biotot_region_japan_mil/bmpop_region_japan$totpop_region_japan_mil
bmpop_region_japan %>% full_join(bmpop_region, by="region") %>% mutate(pct_savebm = 100-biotot_region_japan_mil/biotot_region_mil*100, pct_saveavg = 100-averagebodymass_japan/averagebodymass*100)

```

##### Total biomass and avg body mass for whole world under S1 - and difference from current world

```{r}
#BIOMASS AND AVERAGE BODY MASS FOR WORLD UNDER S1 - AND DIFFERENCE FROM CURRENT
bmpop_world_japan<-DF_ADULT_bayesmod_japan %>% summarise(biotot_world_japan_mil=sum(`22biotot_JAPAN`,na.rm=T)/1000, totpop_world_mil=sum(total_population,na.rm=T)/1000)
bmpop_world_japan$averagebodymass_japan<-bmpop_world_japan$biotot_world_japan_mil/bmpop_world_japan$totpop_world_mil
jp<-bmpop_world_japan %>% full_join(bmpop_world, by="totpop_world_mil") %>% mutate(pct_savebm = 100-biotot_world_japan_mil/biotot_world_mil*100, pct_saveavg = 100-averagebodymass_japan/averagebodymass*100)
jp
avgadultmass_optimal<-jp$averagebodymass_japan[[1]]
```

#### Worst-case scenario

##### Apply BMI distribution of USA

```{r}
#USA SCENARIO...
USA<-DF_ADULT_bayesmod %>% filter(name=="United States of America") %>% select(c(heightagegroups,SEX,BMIMEAN_hns))
USA_re<-USA %>% rename(BMIMEAN_hns_USA=BMIMEAN_hns)

DF_ADULT_bayesmod_USA<-DF_ADULT_bayesmod %>% left_join(USA_re,by=c("heightagegroups","SEX"))
```

##### Biomass estimates for each age-sex-country group under S2

```{r}
#TOTAL BIOMASS BY EACH AGE-SEX-COUNTRY GROUP WITH BMI DIST OF JAPAN
DF_ADULT_bayesmod_USA[,paste(22,"biotot_USA",sep="")] <- DF_ADULT_bayesmod_USA$total_population*DF_ADULT_bayesmod_USA$BMIMEAN_hns_USA*((DF_ADULT_bayesmod_USA$meanheight/100)^2 + (DF_ADULT_bayesmod_USA$sdheight/100)^2) #see formula under table 1 for E(W)
```

##### Total biomass and avg body mass by WHO region under S2 - and difference from current world

```{r}
#BIOMASS AND AVERAGE BODY MASS BY WHO REGION UNDER S1 - AND DIFFERENCE FROM CURRENT
bmpop_region_USA<-DF_ADULT_bayesmod_USA %>% group_by(region) %>% summarise(biotot_region_USA_mil=sum(`22biotot_USA`,na.rm=T)/1000, totpop_region_USA_mil=sum(total_population,na.rm=T)/1000)
bmpop_region_USA$averagebodymass_USA<-bmpop_region_USA$biotot_region_USA_mil/bmpop_region_USA$totpop_region_USA_mil
bmpop_region_USA %>% full_join(bmpop_region, by="region") %>% mutate(pct_addbm =biotot_region_USA_mil/biotot_region_mil*100-100, pct_addavg = averagebodymass_USA/averagebodymass*100-100)

```

##### Total biomass and avg body mass for whole world under S2 - and difference from current world

```{r}
#BIOMASS AND AVERAGE BODY MASS FOR WORLD UNDER S1 - AND DIFFERENCE FROM CURRENT
bmpop_world_USA<-DF_ADULT_bayesmod_USA %>% summarise(biotot_world_USA_mil=sum(`22biotot_USA`,na.rm=T)/1000, totpop_world_mil=sum(total_population,na.rm=T)/1000)
bmpop_world_USA$averagebodymass_USA<-bmpop_world_USA$biotot_world_USA_mil/bmpop_world_USA$totpop_world_mil
bmpop_world_USA %>% full_join(bmpop_world, by="totpop_world_mil") %>% mutate(pct_add = biotot_world_USA_mil/biotot_world_mil*100-100, pct_addavg = averagebodymass_USA/averagebodymass*100-100)

avgadultmass_worstcase<-bmpop_world_USA$averagebodymass_USA[[1]]
```

##### CURRENT EXCESS

```{r}
bmpop_region_japan %>% full_join(bmpop_region, by="region") %>% mutate(savebm = biotot_region_mil-biotot_region_japan_mil, savebm_adults = savebm*1000/avgadultmass_optimal)
bmpop_world_japan$world<-"Y"
bmpop_world$world<-"Y"
bmpop_world_japan %>% full_join(bmpop_world, by="world") %>% mutate(savebm = biotot_world_mil-biotot_world_japan_mil, savebm_adults = savebm*1000/avgadultmass_optimal)
```

##### POTENTIAL EXCESS

```{r}
joint_usjapan<-bmpop_region_USA %>% full_join(bmpop_region_japan, by="region")
joint_usjapan %>% full_join(bmpop_region, by="region") %>% mutate(savebm = biotot_region_USA_mil-biotot_region_japan_mil, savebm_adults = savebm*1000/avgadultmass_optimal)
bmpop_world_USA$world<-"Y"
joint_usjapan_world<-bmpop_world_USA %>% full_join(bmpop_world_japan, by="world")
joint_usjapan_world %>% full_join(bmpop_world, by="world") %>% mutate(savebm = biotot_world_USA_mil-biotot_world_japan_mil, savebm_adults = savebm*1000/avgadultmass_optimal)
```

### Table 3: Energy requirements

#### Load TEE coefficients for calculations

```{r}
TEE_coefficients <- read_excel("~/OneDrive/Documents/UAH/Weightofnations/Code/Code_analysis2022/Output/TEE_coefficients.xlsx")
```

#### Average tTEE (Kcal/day/adult)

##### Current

###### By region

```{r}
energy_byregion<-DF_ADULT_bayesmod %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=(BMR_coeffN*TOTPOP+`22biotot`*BMR_coeffBM)*PAL) %>% group_by(region) %>% summarise(energy_avg = weighted.mean(energy/TOTPOP, TOTPOP))
energy_byregion
```

###### Whole world

```{r}
worldavgenergy<-DF_ADULT_bayesmod %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=(BMR_coeffN*TOTPOP+`22biotot`*BMR_coeffBM)*PAL) %>% summarise(energy_avg = weighted.mean(energy/TOTPOP, TOTPOP, na.rm=T))
worldavgenergy$world<-"Y"
worldavgenergy
```

##### Optimal scenario & difference

###### By region under S1 & with difference

```{r}
energy_byregion_japan<-DF_ADULT_bayesmod_japan %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=(BMR_coeffN*TOTPOP+`22biotot_JAPAN`*BMR_coeffBM)*PAL) %>% group_by(region) %>% summarise(energy_avg_japan = weighted.mean(energy/TOTPOP, TOTPOP, na.rm=T))
energy_byregion_japan %>% full_join(energy_byregion, by="region") %>% mutate(energyDIFF = energy_avg_japan-energy_avg, energyDIFFPCT = energyDIFF/energy_avg*100)
```

###### World under S1 & difference

```{r}
energy_world_japan<-DF_ADULT_bayesmod_japan %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=(BMR_coeffN*TOTPOP+`22biotot_JAPAN`*BMR_coeffBM)*PAL) %>% summarise(energy_avg_japan = weighted.mean(energy/TOTPOP, TOTPOP, na.rm=T))
energy_world_japan$world<-"Y"
s1_energy<-energy_world_japan %>% full_join(worldavgenergy, by="world") %>% mutate(energyDIFF = energy_avg_japan-energy_avg, energyDIFFPCT = energyDIFF/energy_avg*100)
s1_energy
#save energy needed to maintain average adult from the optimal scenario:
avgadultenergy_optimal<-s1_energy$energy_avg_japan[[1]]
```

##### Worst-case scenario & difference

###### By region under S2 & with difference

```{r}
energy_byregion_USA<-DF_ADULT_bayesmod_USA %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=(BMR_coeffN*TOTPOP+`22biotot_USA`*BMR_coeffBM)*PAL) %>% group_by(region) %>% summarise(energy_avg_USA = weighted.mean(energy/TOTPOP, TOTPOP, na.rm=T))
energy_byregion_USA %>% full_join(energy_byregion, by="region") %>% mutate(energyDIFF = energy_avg_USA-energy_avg, energyDIFFPCT = energyDIFF/energy_avg*100)
```

###### World under S2 & difference

```{r}
energy_world_USA<-DF_ADULT_bayesmod_USA %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=(BMR_coeffN*TOTPOP+`22biotot_USA`*BMR_coeffBM)*PAL) %>% summarise(energy_avg_USA = weighted.mean(energy/TOTPOP, TOTPOP, na.rm=T))
energy_world_USA$world<-"Y"
energy_world_USA %>% full_join(worldavgenergy, by="world") %>% mutate(energyDIFF = energy_avg_USA-energy_avg, energyDIFFPCT = energyDIFF/energy_avg*100)
```

##### 

#### Thousands of adults that could be maintained by energy required to maintain:

##### Current excess

```{r}
#No. adults (typical of japan scenario) BY REGION
DF_ADULT_bayesmod_japan %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=((`22biotot`-`22biotot_JAPAN`)*BMR_coeffBM)*PAL) %>% group_by(region) %>% summarise(energy_diff = sum(energy)/avgadultenergy_optimal)
#biomass is in thousands so therefore the number of adults is in terms of thousands of adults

#No. adults (typical of japan scenario) WHOLE WORLD
DF_ADULT_bayesmod_japan %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=((`22biotot`-`22biotot_JAPAN`)*BMR_coeffBM)*PAL) %>% summarise(energy_diff = sum(energy)/avgadultenergy_optimal)
#biomass is in thousands so therefore the number of adults is in terms of thousands of adults
```

##### Potential excess

```{r}
#join usa and japan data as going to analyse "POTENTIAL EXCESS" (USA-JAPAN SCENARIOS)
usa<-DF_ADULT_bayesmod_USA %>% dplyr::select(c("name","SEX","heightagegroups","region","22biotot_USA"))
japan<-DF_ADULT_bayesmod_japan %>% dplyr::select(c("name","SEX","heightagegroups","region","22biotot_JAPAN"))

japan_usa_join<-usa %>% full_join(japan, by=c("name","SEX","heightagegroups","region"))

#No. adults (typical of japan scenario) BY REGION
japan_usa_join %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=((`22biotot_USA`-`22biotot_JAPAN`)*BMR_coeffBM)*PAL) %>% group_by(region) %>% summarise(energy_diff = sum(energy)/avgadultenergy_optimal)
#biomass is in thousands so therefore the number of adults is in terms of thousands of adults

#No. adults (typical of japan scenario) WHOLE WORLD
japan_usa_join %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=((`22biotot_USA`-`22biotot_JAPAN`)*BMR_coeffBM)*PAL) %>% summarise(energy_diff = sum(energy)/avgadultenergy_optimal)
#biomass is in thousands so therefore the number of adults is in terms of thousands of adults
```

### Table 4: CO2 emissions for commercial flights

#### Current

##### CO2 per passenger - REGION

```{r}
DF_ADULT_bayesmod %>% group_by(region) %>% summarise(biotot_region_mil=sum(`22biotot`,na.rm=T)/1000, totpop_region_mil=sum(total_population,na.rm=T)/1000, avgmass = biotot_region_mil/totpop_region_mil, onepassenger = avgmass*0.0945)
```

##### CO2 per passenger - WORLD

```{r}
DF_ADULT_bayesmod %>% summarise(biotot_region_mil=sum(`22biotot`,na.rm=T)/1000, totpop_region_mil=sum(total_population,na.rm=T)/1000, avgmass = biotot_region_mil/totpop_region_mil, onepassenger = avgmass*0.0945)
```

#### Optimal scenario

##### CO2 per passenger under S1 & differences - REGION

```{r}
#excess co2 - japan
underjapan_co2<-DF_ADULT_bayesmod_japan %>% group_by(region) %>% summarise(biomass_underjapanscenario=sum(`22biotot_JAPAN`,na.rm=T)/1000, totpop_region_mil = sum(total_population,na.rm=T)/1000, avgmass_japan = biomass_underjapanscenario/totpop_region_mil, onepassenger_japan = avgmass_japan*0.0945)
current_co2<-DF_ADULT_bayesmod %>% group_by(region) %>% summarise(biotot_region_mil=sum(`22biotot`,na.rm=T)/1000, totpop_region_mil=sum(total_population,na.rm=T)/1000, avgmass = biotot_region_mil/totpop_region_mil, onepassenger = avgmass*0.0945)
underjapan_co2 %>% left_join(current_co2, by="region") %>% mutate(diff_passenger = onepassenger_japan-onepassenger, diff_pct_passenger = diff_passenger/onepassenger*100) %>% dplyr::select(c(region, onepassenger_japan,diff_passenger, diff_pct_passenger))
```

##### CO2 per passenger under S1 & differences - WORLD

```{r}
#excess co2 - japan
underjapan_co2_world<-DF_ADULT_bayesmod_japan %>% summarise(biomass_underjapanscenario=sum(`22biotot_JAPAN`,na.rm=T)/1000, totpop_region_mil = sum(total_population,na.rm=T)/1000, avgmass_japan = biomass_underjapanscenario/totpop_region_mil, onepassenger_japan = avgmass_japan*0.0945)
underjapan_co2_world$world<-"Y"
current_co2_world<-DF_ADULT_bayesmod %>% summarise(biotot_region_mil=sum(`22biotot`,na.rm=T)/1000, totpop_region_mil=sum(total_population,na.rm=T)/1000, avgmass = biotot_region_mil/totpop_region_mil, onepassenger = avgmass*0.0945)
current_co2_world$world<-"Y"
underjapan_co2_world %>% left_join(current_co2_world, by="world") %>% mutate(diff_passenger = onepassenger_japan-onepassenger, diff_pct_passenger = diff_passenger/onepassenger*100) %>% dplyr::select(c(onepassenger_japan,diff_passenger, diff_pct_passenger))
```

#### Worst-case scenario

##### CO2 per passenger under S2 & differences - REGION

```{r}
#excess co2 - USA
underUSA_co2<-DF_ADULT_bayesmod_USA %>% group_by(region) %>% summarise(biomass_underUSAscenario=sum(`22biotot_USA`,na.rm=T)/1000, totpop_region_mil = sum(total_population,na.rm=T)/1000, avgmass_USA = biomass_underUSAscenario/totpop_region_mil, onepassenger_USA = avgmass_USA*0.0945)
underUSA_co2 %>% left_join(current_co2, by="region") %>% mutate(diff_passenger = onepassenger_USA-onepassenger, diff_pct_passenger = diff_passenger/onepassenger*100) %>% dplyr::select(c(region, onepassenger_USA,diff_passenger, diff_pct_passenger))
```

##### CO2 per passenger under S2 & differences - WORLD

```{r}
#excess co2 - japan
underUSA_co2_world<-DF_ADULT_bayesmod_USA %>% summarise(biomass_underUSAscenario=sum(`22biotot_USA`,na.rm=T)/1000, totpop_region_mil = sum(total_population,na.rm=T)/1000, avgmass_USA = biomass_underUSAscenario/totpop_region_mil, onepassenger_USA = avgmass_USA*0.0945)
underUSA_co2_world$world<-"Y"
underUSA_co2_world %>% left_join(current_co2_world, by="world") %>% mutate(diff_passenger = onepassenger_USA-onepassenger, diff_pct_passenger = diff_passenger/onepassenger*100) %>% dplyr::select(c(onepassenger_USA,diff_passenger, diff_pct_passenger))
```

#### 

## RESULTS IN TEXT

### Section: Biomass estimation for 2022

#### US: Body mass

```{r}
DF_ADULT_bayesmod %>% filter(name=="United States of America") %>% summarise(avgmass_US=sum(`22biotot`)/sum(TOTPOP))
```

#### 1 tonne of biomass claculations

```{r}
DF_ADULT_bayesmod %>% group_by(region) %>% summarise(pplpertonne=1000/(sum(`22biotot`)/sum(TOTPOP)))
```

### Section: Potential future scenarios

#### Mean BMI in US and Japan

```{r}
DF_ADULT_bayesmod %>% filter(name %in% c("Japan", "United States of America")) %>% group_by(name) %>% summarise(sum(BMIMEAN_hns)/n())

```

#### Decrease & Increase % in average mass

```{r}
avgadultmass_worstcase/bmpop_world$averagebodymass[[1]]*100-100

bmpop_world$averagebodymass[[1]]/avgadultmass_optimal*100
```

#### Excess -regional (comment - as not sure if going to use!)

```{r}
bmpop_region_japan %>% full_join(bmpop_region, by="region") %>% mutate(pct_savebm = biotot_region_mil/biotot_region_japan_mil*100-100)
```

#### Increase in human biomass under worst-case scenario compared to current scenario (biomass & number of adults)

```{r}
bmpop_world_USA %>% full_join(bmpop_world, by="totpop_world_mil") %>% mutate(pctbm_add = biotot_world_USA_mil/biotot_world_mil*100-100, bm_add = biotot_world_USA_mil-biotot_world_mil) %>% dplyr::select(c("pctbm_add","bm_add")) %>% mutate(bm_add/avgadultmass_optimal)

```

#### Increase in human biomass under worst-case scenario compared to optimal scenario (biomass & number of adults)

```{r}
d3<-bmpop_world_japan %>% full_join(bmpop_world, by="world") %>% mutate(savebm = biotot_world_mil-biotot_world_japan_mil, savebm_adults = savebm*1000/avgadultmass_optimal)

d4<-bmpop_world_USA %>% full_join(bmpop_world_japan, by="totpop_world_mil") %>% mutate(pctbm_add = biotot_world_USA_mil/biotot_world_japan_mil*100-100, bm_add = biotot_world_USA_mil-biotot_world_japan_mil) %>% dplyr::select(c("pctbm_add","bm_add")) %>% mutate(bm_add/avgadultmass_optimal)
d4
d4$bm_add[[1]]/d3$savebm[[1]]
```

#### World pop vs current excess

```{r}
#population proportions
globalpop<-sum(bmpop_region_japan$totpop_region_japan_mil)
bmpop_region_japan %>% group_by(region) %>% summarise(prop_pop = totpop_region_japan_mil/globalpop)

#excess biomass
bmpop_region_japan %>% full_join(bmpop_region, by="region") %>% mutate(savebm = biotot_region_mil-biotot_region_japan_mil, savebm_pcttot = savebm/sum(savebm))

```

#### Does biomass increase for all countries current vs under optimal case scenario?

```{r}
dt<-DF_ADULT_bayesmod_japan %>%  group_by(name, region) %>% summarise(biomass_current = sum(`22biotot`), biomass_opt = sum(`22biotot_JAPAN`)) %>% mutate(diff=biomass_current-biomass_opt)
mean(dt$diff>0)
sameorless<-dt %>% filter(diff<=0) 
dt %>% filter(diff<=0) %>% group_by(region) %>% summarise(n=n(), pct = n()/nrow(sameorless))
```

#### China, USA and India stats under optimal

```{r}
bmpop_country_japan<-DF_ADULT_bayesmod_japan %>% filter(name %in% c("China","United States of America","India")) %>% group_by(name) %>% summarise(biotot_country_japan_mil=sum(`22biotot_JAPAN`,na.rm=T)/1000, totpop_country_japan_mil=sum(total_population,na.rm=T)/1000)
bmpop_country<-DF_ADULT_bayesmod %>% filter(name %in% c("China","United States of America","India")) %>% group_by(name) %>% summarise(biotot_country_mil=sum(`22biotot`,na.rm=T)/1000, totpop_country_mil=sum(total_population,na.rm=T)/1000)
bmpop_country_japan %>% full_join(bmpop_country, by="name") %>% mutate(savebm = biotot_country_mil-biotot_country_japan_mil, savebm_pcttot = savebm/sum(savebm), savebm_avgadults = savebm/avgadultmass_optimal)
```

#### Which countries have highest proportion of current biomass in excess?

```{r}
excess_tot<-DF_ADULT_bayesmod_japan %>% summarise(excess = sum(`22biotot`)-sum(`22biotot_JAPAN`))
DF_ADULT_bayesmod_japan %>% group_by(name) %>% summarise(excess = (sum(`22biotot`)-sum(`22biotot_JAPAN`)), excess_pct = (sum(`22biotot`)-sum(`22biotot_JAPAN`))/sum(`22biotot`)*100) %>% arrange(-excess_pct)

```

#### Which countries have excess biomass greater than 1 million tonnes?

```{r}
DF_ADULT_bayesmod_japan %>% group_by(name) %>% summarise(excess = (sum(`22biotot`)-sum(`22biotot_JAPAN`))/1000) %>% arrange(-excess)
```

#### Does biomass decrease for all countries under worst case scenario vs current?

```{r}
dt2<-DF_ADULT_bayesmod_USA %>%  group_by(name, region) %>% summarise(biomass_current = sum(`22biotot`), biomass_worst = sum(`22biotot_USA`)) %>% mutate(diff=biomass_worst-biomass_current)
mean(dt2$diff>0)
```

#### China and India stats under worst case

```{r}
#vs current
bmpop_country_USA<-DF_ADULT_bayesmod_USA %>% filter(name %in% c("China","India")) %>% group_by(name) %>% summarise(biotot_country_USA_mil=sum(`22biotot_USA`,na.rm=T)/1000, totpop_country_USA_mil=sum(total_population,na.rm=T)/1000)
bmpop_country<-DF_ADULT_bayesmod %>% filter(name %in% c("China","India")) %>% group_by(name) %>% summarise(biotot_country_mil=sum(`22biotot`,na.rm=T)/1000, totpop_country_mil=sum(total_population,na.rm=T)/1000)
bmpop_country_USA %>% full_join(bmpop_country, by="name") %>% mutate(savebm = biotot_country_USA_mil-biotot_country_mil)

#vs optimal
bmpop_country_USA %>% full_join(bmpop_country_japan, by="name") %>% mutate(savebm = biotot_country_USA_mil-biotot_country_japan_mil)

bmpop_country_USA %>% full_join(bmpop_country_japan, by="name") %>% mutate(savebm_avgadult = (biotot_country_USA_mil-biotot_country_japan_mil)/avgadultmass_optimal)
```

### Section: Consequences

#### Energy required to maintain excess biomass in US alone

```{r}
#No. adults (typical of japan scenario) BY REGION
DF_ADULT_bayesmod_japan %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=((`22biotot`-`22biotot_JAPAN`)*BMR_coeffBM)*PAL) %>% group_by(name) %>% filter(name=="United States of America") %>% summarise(energy_diff = sum(energy)/avgadultenergy_optimal)
#biomass is in thousands so therefore the number of adults is in terms of thousands of adults
```

```         
```

#### Energy required to sustain worst-case vs current (% and number of adults)

```{r}
DF_ADULT_bayesmod_USA %>% full_join(TEE_coefficients, by=c("heightagegroups", "SEX")) %>% mutate(energy=((`22biotot_USA`-`22biotot`)*BMR_coeffBM)*PAL) %>% summarise(energy_diff = sum(energy)/avgadultenergy_optimal)
#biomass is in thousands so therefore the number of adults is in terms of thousands of adults

```

#### Average emissions: % current vs optimal

```{r}
current_co2_world %>% left_join(underjapan_co2_world, by="world") %>% mutate(diff_passenger = onepassenger-onepassenger_japan, diff_pct_passenger = diff_passenger/onepassenger_japan*100) %>% dplyr::select(c(onepassenger,diff_passenger, diff_pct_passenger))
```

#### Average emissions: % worst-case vs optimal

```{r}
underUSA_co2_world %>% left_join(underjapan_co2_world, by="world") %>% mutate(diff_passenger = onepassenger_USA-onepassenger_japan, diff_pct_passenger = diff_passenger/onepassenger_japan*100) %>% dplyr::select(c(onepassenger_USA,diff_passenger, diff_pct_passenger))
```

### Section: Abstract

#### Commercial flight: North America vs Asia

```{r}
current_co2[current_co2$region=="NORTHERN AM.","onepassenger"][[1]]/current_co2[current_co2$region=="ASIA","onepassenger"][[1]]
```

## FIGURES

### Maps

#### Load packages and set up data for maps

```{r}
library(cartogram) 
library(sf) 
library(tmap) 
library(tibble)
data("World")
World<-World[World$continent!="Antarctica",]
World <- st_transform(World, 3395)
dta_2022<-DF_ADULT_bayesmod_japan %>% full_join(DF_ADULT_bayesmod_USA)
country_lookup <- tribble(
  ~Worldmap,                                ~biomassnames,
  "Bosnia and Herz.", "Bosnia and Herzegovina", 
  "Brunei", "Brunei Darussalam",
  "Central African Rep.", "Central African Republic",
  "Cote d'Ivoire", "C\x99te d'Ivoire",
  "Czechia", "Czech Republic",
  "Dem. Rep. Congo", "Democratic Republic of the Congo",
  "Dominican Rep.", "Dominican Republic",
  "Eq. Guinea", "Equatorial Guinea",
  "Iran", "Iran (Islamic Republic of)",
  "Laos", "Lao People's Democratic Republic",
  "Libya", "Libyan Arab Jamahiriya",
  "Moldova", "Republic of Moldova", 
  "North Korea", "Dem. People's Republic of Korea",
  "North Macedonia", "TFYR Macedonia",
  "Palestine", "Occupied Palestinian Territory",
  "Russia","Russian Federation",
  "S. Sudan","Sudan",
  "Solomon Is.","Solomon Islands",
  "South Korea","Republic of Korea",
  "Syria","Syrian Arab Republic",
  "Tanzania", "United Republic of Tanzania",
  "Venezuela", "Venezuela (Bolivarian Republic of)",
  "Vietnam", "Viet Nam",
  "W. Sahara", "Western Sahara",
  "eSwatini", "Swaziland"
)
dta_2022_namesupdated <- dta_2022 %>%
  left_join(country_lookup, by = c("name" = "biomassnames")) %>%
  mutate(name = if_else(!is.na(Worldmap), Worldmap, name)) %>%
  dplyr::select(-Worldmap)
```

#### Figure 3:

```{r}
#Data for Population & biomass plots
globalbiomass<-sum(dta_2022_namesupdated$`22biotot`)
globalpop<-sum(dta_2022_namesupdated$TOTPOP)
biomasstot_country<-dta_2022_namesupdated %>% group_by(name) %>% summarise(biotot_region_pct=sum(`22biotot`,na.rm=T)/globalbiomass*100, totpop_region_mil=sum(total_population,na.rm=T)/1000)
World_biomasstot<-biomasstot_country %>% full_join(World, by=c("name"))
World_biomasstot[is.na(World_biomasstot$biotot_region_pct), ]
World_biomasstot_clean2 <- World_biomasstot %>% filter(!is.na(area))
World_biomasstot_clean3 <- st_as_sf(World_biomasstot_clean2)
  
#Population 
globalpop<-sum(dta_2022_namesupdated$TOTPOP)
World_totpop_cont <- cartogram_cont(World_biomasstot_clean3, "totpop_region_mil", itermax = 5)
tot_pop_plot<-tm_shape(World_totpop_cont) + tm_polygons("totpop_region_mil", style = "jenks", title="Adult pop (millions)") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom")) 
tot_pop_plot

#Current biomass
World_biomasstot_cont <- cartogram_cont(World_biomasstot_clean3, "biotot_region_pct", itermax = 5)
tot_biomass_plot<-tm_shape(World_biomasstot_cont) + tm_polygons("biotot_region_pct", style = "jenks", title="% of global biomass") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom")) 
tot_biomass_plot

#Current:optimal ratio
biomasstot_country_ctoo<-dta_2022_namesupdated %>% group_by(name) %>% summarise(biotot_region_pct_currenttooptratio=sum(`22biotot`,na.rm=T)/sum(`22biotot_JAPAN`,na.rm=T))
World_biomasstot_ctoo<-biomasstot_country_ctoo %>% full_join(World, by=c("name"))
World_biomasstot_ctoo[is.na(World_biomasstot_ctoo$biotot_region_pct_currenttooptratio), ]
World_biomasstot_ctoo_clean2 <- World_biomasstot_ctoo %>% filter(!is.na(area))
World_biomasstot_ctoo_clean3 <- st_as_sf(World_biomasstot_ctoo_clean2)
World_biomasstot_ctoo_cont <- cartogram_cont(World_biomasstot_ctoo_clean3, "biotot_region_pct_currenttooptratio", itermax = 5)
tot_biomass_plot_currenttoopt<-tm_shape(World_biomasstot_ctoo_cont) + tm_polygons("biotot_region_pct_currenttooptratio", style = "jenks", title="Current biomass : Lowest case biomass") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom")) 
tot_biomass_plot_currenttoopt

#Worst-case:current ratio
biomasstot_country_wctoo<-dta_2022_namesupdated %>% group_by(name) %>% summarise(biotot_region_pct_worsttooptratio=sum(`22biotot_USA`,na.rm=T)/sum(`22biotot`,na.rm=T))
World_biomasstot_wctoo<-biomasstot_country_wctoo %>% full_join(World, by=c("name"))
World_biomasstot_wctoo[is.na(World_biomasstot_wctoo$biotot_region_pct_worsttooptratio), ]
World_biomasstot_wctoo_clean2 <- World_biomasstot_wctoo %>% filter(!is.na(area))
World_biomasstot_wctoo_clean3 <- st_as_sf(World_biomasstot_wctoo_clean2)
World_biomasstot_wctoo_cont <- cartogram_cont(World_biomasstot_wctoo_clean3, "biotot_region_pct_worsttooptratio", itermax = 5)
tot_biomass_plot_worsttoopt<-tm_shape(World_biomasstot_wctoo_cont) + tm_polygons("biotot_region_pct_worsttooptratio", style = "jenks", title="Highest case biomass : Current biomass") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom")) 
tot_biomass_plot_worsttoopt

```

```{r}

###FORMATTING PLOTS FOR OUTPUT

tmap_mode("plot")

# Define layout settings
common_layout <- list(
    legend.outside = TRUE,
    legend.outside.position = "right",
    legend.text.size = 0.6,
    legend.title.size = 0.8,
    legend.bg.color = "white",
    legend.frame = TRUE,
    frame = TRUE,
    inner.margins = c(0, 0, 0, 0),
    outer.margins = c(0, 0, 0, 0),
    asp = 0
)


# Create each map using do.call()
map1 <- tm_shape(World_totpop_cont) +
    tm_polygons("totpop_region_mil", style = "jenks", title = "Adult population   \n(millions)") +
    do.call(tm_layout, common_layout)

map2 <- tm_shape(World_biomasstot_cont) +
    tm_polygons("biotot_region_pct", style = "jenks", title = "% global biomass ") +
    do.call(tm_layout, common_layout)

map3 <- tm_shape(World_biomasstot_ctoo_cont) +
    tm_polygons("biotot_region_pct_currenttooptratio", style = "jenks", title = "Current biomass :\nLowest biomass") +
    do.call(tm_layout, common_layout)

map4 <- tm_shape(World_biomasstot_wctoo_cont) +
    tm_polygons("biotot_region_pct_worsttooptratio", style = "jenks", title = "Highest     \nbiomass :\nCurrent biomass   ") +
    do.call(tm_layout, common_layout)
  
# Arrange the four maps in a 3x2 grid
p1<-tmap_arrange(
    map1, map2, map3, map4,
    nrow = 2,
    ncol = 2,
    outer.margins = 0.02
)

print(p1)

png("Output/Grid_fourcartograms_jun17.png",width=12,height=6, units = "in", res = 300)
print(p1)
dev.off()
```

#### Figure 4: Chloropleths

```{r}
#Current excess
biomassexcessandtot_country<-dta_2022_namesupdated %>% group_by(name) %>% summarise(biomassexcess_country=sum(`22biotot`-`22biotot_JAPAN`,na.rm=T)/1000000, biomasspotentialexcess_country=sum(`22biotot_USA`-`22biotot_JAPAN`,na.rm=T)/1000000)
biomassexcessandtot_country_namesupdate <- biomassexcessandtot_country %>%
  left_join(country_lookup, by = c("name" = "biomassnames")) %>%
  mutate(Country = if_else(!is.na(Worldmap), Worldmap, name)) %>%
  select(-Worldmap)
World_biomasstoexcessandtot<-biomassexcessandtot_country_namesupdate %>% full_join(World, by=c("Country"="name"))
World_biomasstoexcessandtot[is.na(World_biomasstoexcessandtot$biomassexcess_country), ]
World_biomasstoexcessandtot_clean <- World_biomasstoexcessandtot %>% filter(!is.na(area))
World_biomasstoexcessandtot_clean2 <- st_as_sf(World_biomasstoexcessandtot_clean)

breaks_shared <- c(-1, -0.5, 0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 10, 15, 20, 25)

biomass_currentexcess_plot<-tm_shape(World_biomasstoexcessandtot_clean2) + tm_polygons("biomassexcess_country", title="Current excess biomass (MT)", breaks=breaks_shared, palette = "-RdYlGn", midpoint=0) + tm_layout(frame = FALSE, legend.position = c("left", "bottom"))
biomass_currentexcess_plot

#Potential excess
biomass_potentialexcess_plot<-tm_shape(World_biomasstoexcessandtot_clean2) + tm_polygons("biomasspotentialexcess_country", title="Potential excess biomass (MT)", breaks=breaks_shared, palette = "-RdYlGn", midpoint=0) + tm_layout(frame = FALSE, legend.position = c("left", "bottom"))
biomass_potentialexcess_plot


biomass_currentexcess_plot <- tm_shape(World_biomasstoexcessandtot_clean2) +
  tm_polygons(
    "biomassexcess_country",
    title = "Current excess biomass (MT)",
    breaks = breaks_shared,
    palette = "-RdYlGn",
    midpoint = 0
  ) +
  tm_layout(frame = TRUE, legend.show = FALSE)

biomass_potentialexcess_plot <- tm_shape(World_biomasstoexcessandtot_clean2) +
  tm_polygons(
    "biomasspotentialexcess_country",
    title = "Potential excess biomass (MT)",
    breaks = breaks_shared,
    palette = "-RdYlGn",
    midpoint = 0
  ) +
  tm_layout(frame = TRUE, legend.show = FALSE)

legend_plot <- tm_shape(World_biomasstoexcessandtot_clean2) +
  tm_polygons(
    "biomassexcess_country",
    breaks = breaks_shared,
    palette = "-RdYlGn",
    midpoint = 0,
    title = "Excess biomass\n(million tonnes)"
  ) +
  tm_layout(
    legend.only = TRUE,
    legend.position = c("left", "center"),
    frame = TRUE
  )

p2<-tmap_arrange(
  biomass_currentexcess_plot,
  legend_plot,
  biomass_potentialexcess_plot,
  ncol = 2,
  widths = c(3, 1)
)

png("Output/Grid_twochloropleths_apr17.png",width=5,height=5, units = "in", res = 300)
print(p2)
dev.off()


```

### Bar charts

#### Figure 5: Change in biomass under optimal and worst case scenarios BY REGION & WB STATUS

```{r}
merged2<-DF_ADULT_bayesmod_japan %>% full_join(DF_ADULT_bayesmod_USA)
library(stringr)
library(tidyr)
WBstatus <- read_excel("~/OneDrive/Documents/UAH/Weightofnations/Data sources/WorldBank/OGHIST_2026_03_10_2022dataonly copy_matchnames.xlsx", 
    sheet = "Country Analytical History")
WBstatus2<-WBstatus %>% mutate(WB_name=if_else(WB_name=="Cote divoire", "C\x99te d'Ivoire", WB_name))

merged2_clean <- merged2 %>%
  mutate(name_clean = str_replace_all(name, "\\s+", "") %>% str_to_lower())

WBstatus_clean <- WBstatus2 %>%
  mutate(WB_name_clean = str_replace_all(WB_name, "\\s+", "") %>% str_to_lower())

merged2_WB<-merged2_clean %>%
  full_join(WBstatus_clean, by = c("name_clean" = "WB_name_clean"))

data_diff<-merged2_WB %>% filter(!is.na(region)) %>% dplyr::select(region, name, heightagegroups,SEX, TOTPOP, `22biotot`, `22biotot_JAPAN`, `22biotot_USA`, WB_2022)
data_diff$diff_japan<-data_diff$`22biotot_JAPAN`-data_diff$`22biotot`
data_diff$diff_USA<-data_diff$`22biotot_USA`-data_diff$`22biotot`

worldbiomass<-merged2 %>% summarise(biomass_world=sum(`22biotot`,na.rm=T)/1000)

figdata_change_region<-data_diff %>% filter(!is.na(region)) %>% group_by(region) %>% summarise(biomassd_japan = sum(diff_japan, na.rm=T)/sum(`22biotot`)*100, biomassd_usa = sum(diff_USA, na.rm=T)/sum(`22biotot`)*100, biomassd_japan_ab = sum(diff_japan, na.rm=T), biomassd_usa_ab = sum(diff_USA, na.rm=T) ) %>% pivot_longer(
    cols = c(biomassd_japan, biomassd_usa, biomassd_japan_ab, biomassd_usa_ab),
    names_to = "scenario",
    values_to = "biomass_change"
  ) %>% 
  mutate(
    scenario = recode(scenario,
                      biomassd_japan = "Lowest scenario",
                      biomassd_usa   = "Highest scenario"),
    WB_2022 = "Overall" 
  ) 


figdata_change_region_WBstatus<-data_diff %>% filter(!is.na(WB_2022) & !is.na(region)) %>% group_by(region, WB_2022) %>% summarise(biomassd_japan = sum(diff_japan, na.rm=T)/sum(`22biotot`)*100, biomassd_usa = sum(diff_USA, na.rm=T)/sum(`22biotot`)*100,  biomassd_japan_ab = sum(diff_japan, na.rm=T), biomassd_usa_ab = sum(diff_USA, na.rm=T)) %>% pivot_longer(
    cols = c(biomassd_japan, biomassd_usa, biomassd_japan_ab, biomassd_usa_ab),
    names_to = "scenario",
    values_to = "biomass_change"
  ) %>% 
  mutate(
    scenario = recode(scenario,
                      biomassd_japan = "Lowest scenario",
                      biomassd_usa   = "Highest scenario")
  ) 


figdata_change<-rbind(figdata_change_region,figdata_change_region_WBstatus)
# country_order <- figdata_change %>%
#   distinct(name, biomass_pct) %>%
#   arrange(biomass_pct) %>%
#   pull(name)



figdata_change$WB_2022 <- factor(
  figdata_change$WB_2022,
  levels = c("Overall", "H", "UM", "LM", "L"),
  labels = c("Overall", "High", "Upper-middle", "Lower-middle", "Low")
)



figdata_change_pct <- figdata_change %>% filter(scenario %in% c(
    "Lowest scenario",
    "Highest scenario"
  ))


figdata_change_pct$scenario <- factor(
  figdata_change_pct$scenario,
  levels = c(
    "Lowest scenario",
    "Highest scenario"
  )
)

figdata_change_pct <- figdata_change_pct[order(figdata_change_pct$scenario == "Lowest scenario"), ]

figdata_less<-figdata_change%>% filter(scenario=="biomassd_japan_ab") %>% mutate(vbl = ifelse(biomass_change<0,paste0(sprintf("%.1f", biomass_change / 1000000), " MT"), paste0("+",sprintf("%.1f", biomass_change / 1000000), " MT")))
figdata_more<-figdata_change%>% filter(scenario=="biomassd_usa_ab") %>% mutate(vbl = paste0("+",sprintf("%.1f", biomass_change / 1000000), " MT"))


fig_biomass_change<-ggplot(figdata_change_pct, aes(x=biomass_change, y=WB_2022, fill=as.factor(scenario))) + 
  facet_wrap(~region)+
  geom_col(width=0.8, position="identity")+theme_bw()+
  labs(
    x = "Change in biomass (%)",
    y = NULL
  )+
  guides(fill = guide_legend(title = NULL,nrow=1,byrow=F))+
  theme(
    legend.position = "bottom",
    legend.background = element_rect(
      fill = alpha("white", 0.5),  # semi-transparent fill
      color = "black",             # border color
      size = 0.8                   # border thickness
    ),    
    text = element_text(color = "black"),          # general text
    axis.text = element_text(color = "black", face="bold"),     # axis tick labels
    axis.title = element_text(color = "black", face="bold"),    # axis titles
    plot.subtitle = element_text(color = "black"), # subtitle
    legend.text = element_text(color = "black", face="bold"),   # legend labels
    legend.title = element_text(color = "black", face="bold"),   # legend title,
    strip.text = element_text(face = "bold", size=11),
    
  )+
  scale_fill_manual(values=c("Lowest scenario"="#FFB000", "Highest scenario"="#648FFF")) + 
  geom_vline(xintercept=0, linetype="dashed", color="gray20") +
  geom_label(data=figdata_less, aes(x=-max(figdata_change_pct$biomass_change)-10, y=WB_2022, label=vbl), fill="grey90", size=3, color="black") +
  geom_label(data=figdata_more, aes(x=max(figdata_change_pct$biomass_change)+10, y=WB_2022, label=vbl), fill="grey90", size=3, color="black") +
  scale_x_continuous(breaks=c(-30,-25,-20,-15,-10,-5,0,5,10,15,20,25,30), labels = c("-30", "", "-20", "", "-10", "", "0", "", "+10", "", "+20", "", "+30"), limits=c(-max(figdata_change_pct$biomass_change)-17, max(figdata_change_pct$biomass_change)+17))
  



ggsave("Output/Fig_2022_change_biomass_17Jun.png",
       plot = fig_biomass_change,
       width = 10,
       height = 4,
       dpi = 300)

data_diff %>% filter(!is.na(WB_2022) & !is.na(region)) %>% filter(heightagegroups=="18-29" & SEX==0) %>% group_by(region, WB_2022) %>% summarise(n())

data_diff %>% filter(!is.na(WB_2022) & !is.na(region)) %>% filter(heightagegroups=="18-29" & SEX==0) %>% group_by(region) %>% summarise(n())

#FOR RESULTS IN TEXT FOR THIS FIGURE
figdata_change_pct %>% filter(region=="AFRICA")
figdata_change%>% filter(scenario=="biomassd_usa_ab") %>% filter(region=="AFRICA" & WB_2022 =="High") %>% mutate(biomass_change/1000000)
```

### Demonstrating message

#### Figure 1: Normal vs split-normal

```{r}

pivot_japan<-22.5
sigma1_japan<-5
sigma2_japan<-10

pivot_US<-22.5
sigma1_US<-5
sigma2_US<-5

splitvsnormal<-ggplot() +
  xlim(c(10,50)) +
  geom_vline(aes(xintercept=22.5), linetype="dotted")+
  stat_function(
    fun = function(x) {
      ifelse(
        x <= pivot_japan,
        sqrt(2/pi) / (sigma1_japan + sigma2_japan) * 
          exp(- (x - pivot_japan)^2 / (2 * sigma1_japan^2)),
        sqrt(2/pi) / (sigma1_japan + sigma2_japan) * 
          exp(- (x - pivot_japan)^2 / (2 * sigma2_japan^2))
      )
    },
    color = "#60A561",
    linewidth = 1.2,
    n = 500
  )  +
  stat_function(
    fun = function(x) {
      ifelse(
        x <= pivot_US,
        sqrt(2/pi) / (sigma1_US + sigma2_US) * 
          exp(- (x - pivot_US)^2 / (2 * sigma1_US^2)),
        sqrt(2/pi) / (sigma1_US + sigma2_US) * 
          exp(- (x - pivot_US)^2 / (2 * sigma2_US^2))
      )
    },
    color = "#4D7EA8",
    linewidth = 1.2, 
    n = 500
  )  +
  labs(
    x = "BMI (kg/m^2)",
    y = "Density"
  ) +
  theme_minimal() +
  ylim(c(0, 0.09)) 

ggsave("Output/Fig_splitvsnormal_17Apr.png",
       plot = splitvsnormal,
       width = 6,
       height = 3.5,
       dpi = 300)
 
```

## Supplementary File 2 & Supplementary Figure 1: explaining excess biomass plot

```{r}
library(truncnorm)
library(ggplot2)

dta_2022<-readRDS("/Users/issylandray/Library/CloudStorage/OneDrive-Personal/Documents/UAH/Weightofnations/Code/Code_analysis2022/Output/WON_analysis_allvbls_2022_Apr15.rds")

results<-dta_2022 %>% filter(name %in% c("Japan", "United States of America") & heightagegroups =="45-59" & SEX==0)

meanbmi_japan<-20
sigma1_japan<-5

meanbmi_US<-28
sigma1_US<-5

xvals<-c(15,20)

qs<-ptruncnorm(xvals, mean = meanbmi_japan,sd = sigma1_japan,a = 10)

japan_q<-qtruncnorm(qs,mean = meanbmi_japan,sd = sigma1_japan,a = 10)

us_q<-qtruncnorm(qs,mean = meanbmi_US,sd = sigma1_US,a = 10)

df_q <- data.frame(
  q = qs,
  x_japan = japan_q,
  x_us = us_q,
  position = -qs/50
)

legend_df <- data.frame(
  group = c("Japan: 30-44 years, male", "US: 30-44 years, male"),
  x = c(10, 10),
  y = c(0, 0)
)


#BMI

bmi_plot<-ggplot() +
  geom_vline(xintercept=25,color="grey30", linetype="dotted") +
  geom_vline(xintercept=30,color="grey30", linetype="dotted") +
  xlim(c(10,50)) +
  stat_function(
    fun = function(x) dnorm(x, mean = meanbmi_japan, sd = sigma1_japan),
    color = "#9E1946",
    linewidth = 1.2,
    n = 500
  ) +
  stat_function(
    fun = function(x) dnorm(x, mean = meanbmi_US, sd = sigma1_US),
    color = "#4D6CFA",
    linewidth = 1.2,
    n = 500
  ) +
  labs(
    x = "BMI (kg/m^2)",
    y = "Density"
  ) +
  theme_minimal() +
  geom_point(
    data = df_q,
    aes(x = x_japan, y = -0.005,shape=as.factor(qs)),
    color = "#9E1946",
    size = 3
  ) +
  geom_point(
    data = df_q,
    aes(x = x_us, y = -0.005,shape=as.factor(qs)),
    color = "#4D6CFA",
    size = 3
  ) +
  geom_curve(
    data = df_q,
    aes(
      x = x_japan+0.3, y = -0.006,
      xend = x_us-0.3, yend = -0.006
    ),
    curvature = 0.35,
    linetype = "dashed",
    arrow = arrow(length = unit(0.2, "cm"), type="closed"),
    color = "grey4"
  )  + ylim(c(-0.013, 0.1)) +
  theme(legend.position = "null")
  

#HEIGHT
mean_height<-170
sd_height<-5

height_plot<-ggplot() +
  xlim(c(120,200)) +
  stat_function(
    fun = function(x) dnorm(x, mean = mean_height, sd = sd_height),
    color = "darkgreen",
    linewidth = 1.2,
    n = 500
  ) +
  labs(
    x = "Height (cm)",
    y = "Density"
  ) +
  theme_minimal()

#Simulate biomass dist
n <- 1000000
height <- rnorm(n, mean_height, sd_height)

bmi_japan <-rnorm(n, mean = meanbmi_japan,sd = sigma1_japan)
mass_japan <- bmi_japan * ((height/100)^2+(sd_height/100)^2)

bmi_US <-rnorm(n, mean = meanbmi_US,sd = sigma1_US)
mass_US <- bmi_US * ((height/100)^2+(sd_height/100)^2)

df <- data.frame(
  mass = c(mass_japan, mass_US),
  country = rep(c("Group 1", "Group 2"),
                times = c(length(mass_japan), length(mass_US)))
)

equivalent_mass <- function(x, jp, us) {
  p <- mean(jp <= x)
  quantile(us, probs = p)
}

equivalent_mass(50, mass_japan, mass_US)
equivalent_mass(75, mass_japan, mass_US)
equivalent_mass(100, mass_japan, mass_US)


df_q$x_japan_mass<-df_q$x_japan*((mean_height/100)^2+(sd_height/100)^2)
df_q$x_US_mass<-df_q$x_us*((mean_height/100)^2+(sd_height/100)^2)

mass_plot<-ggplot(df, aes(x = mass, color = country)) +
  geom_vline(xintercept=25*(1.7^2+0.05^2),color="grey30", linetype="dotted") +
  geom_vline(xintercept=30*(1.7^2+0.05^2),color="grey30", linetype="dotted") +
  ylim(c(-0.003,0.03))+
  geom_density(linewidth=1) +
  labs(x = "Mass (kg)", y = "Density") +
  theme_minimal()  +xlim(c(0, 130))  +
  scale_color_manual(
    name = "Country age-sex group",
    values = c(
      "Group 1" = "#9E1946",
      "Group 2" = "#4D6CFA"
    )
  ) +theme(
    legend.position = "none") +
  geom_point(
    data = df_q,
    aes(x = x_japan_mass, y = -0.0008,shape=as.factor(qs)),
    color = "#9E1946",
    size = 3
  ) +
  geom_point(
    data = df_q,
    aes(x = x_US_mass, y = -0.0008,shape=as.factor(qs)),
    color = "#4D6CFA",
    size = 3
  ) +
  geom_curve(
    data = df_q,
    aes(
      x = x_japan_mass+1, y = -0.0009,
      xend = x_US_mass-1, yend = -0.0009
    ),
    curvature = 0.35,
    linetype = "dashed",
    arrow = arrow(length = unit(0.2, "cm"), type="closed"),
    color = "grey4"
  )  


#CDF for Mass
x_grid <- seq(min(df$mass), max(df$mass), length.out = 1000)
ecdf_japan <- ecdf(mass_japan)
ecdf_us <- ecdf(mass_US)
df_ecdf <- data.frame(
  x = x_grid,
  F_japan = ecdf_japan(x_grid),
  F_us = ecdf_us(x_grid)
)

df_inv <- data.frame(
  y = df_ecdf$F_japan,
  x = df_ecdf$x
)



cdf_plot<-ggplot() +
  geom_line(aes(x=df_ecdf$x,y = df_ecdf$F_japan, color = "Japan: 30-44 years, male"), linewidth = 1.2) +
  geom_line(aes(x=df_ecdf$x,y = df_ecdf$F_us, color = "US: 30-44 years, male"), linewidth = 1.2) +
  scale_color_manual(name = "Country age-sex group",
                     values = c(
                       "Japan: 30-44 years, male" = "#9E1946",
                       "US: 30-44 years, male" = "#4D6CFA"
                     )) +
  
  labs(x = "Mass (kg)", y = "Cumulative density") +geom_point(
    data = df_q,
    aes(x = x_japan_mass-0.6, y = df_q$q,shape=as.factor(qs)), #there is slight error so I have jittered the points as this is just a picture example
    color = "#9E1946",
    size = 4
  ) +
  geom_point(
    data = df_q,
    aes(x = x_US_mass, y = df_q$q,shape=as.factor(qs)),
    color = "#4D6CFA",
    size = 4
  )+theme_minimal() + geom_vline(xintercept=25*1.7^2,color="grey30", linetype="dotted") +
  geom_vline(xintercept=30*1.7^2,color="grey30", linetype="dotted") + xlim(c(0,130))+
  theme(legend.position = "null") +
  geom_segment(
    data = df_q,
    aes(x = x_japan_mass+1, y = q,
        xend = x_US_mass-1.5, yend = q),
    color = "grey4",
    linetype="dashed",
    arrow = arrow(length = unit(0.2, "cm"), type="closed")
  )
  

t25 <- 25 * (1.7^2+0.05^2)
t30 <- 30 * (1.7^2+0.05^2)
df_low  <- df_ecdf[df_ecdf$x <= t25, ]
df_mid  <- df_ecdf[df_ecdf$x > t25 & df_ecdf$x <= t30, ]
df_high <- df_ecdf[df_ecdf$x > t30, ]
x_low  <- mean(df_low$x)
x_mid  <- mean(df_mid$x)
x_high <- mean(df_high$x)
mid_y_between <- function(x0, df) {
  row <- df[which.min(abs(df$x - x0)), ]
  mean(c(row$F_japan, row$F_us))
}

mid_y_above <- function(x0, df) {
  row <- df[which.min(abs(df$x - x0)), ]
  mean(c(row$F_japan, 1))
}
labels_df <- data.frame(
  x = c(x_low, x_mid, x_high, x_low, x_mid, x_high),
  y = c(
    mid_y_above(x_low, df_low),
    mid_y_above(x_mid, df_mid),
    mid_y_above(x_high, df_high),
    mid_y_between(x_low, df_low),
    mid_y_between(x_mid, df_mid),
    mid_y_between(x_high, df_high)
  ),
  label = c("A", "B", "C", "X", "Y", "Z")
)
labels_df$x[labels_df$label == "X"] <- 60
labels_df$y[labels_df$label == "X"] <- 0.25

labels_df$x[labels_df$label == "Z"] <- 95
labels_df$y[labels_df$label == "Z"] <- 0.9

labels_df$y[labels_df$label == "B"] <- 0.97

labels_df$x[labels_df$label == "C"] <- 97
labels_df$y[labels_df$label == "C"] <- 1.05

cdf_plot_areas<-ggplot() +
  geom_ribbon(data = df_low,
              aes(x = x, ymin = F_japan, ymax = 1),
              fill = "#4D6CFA", alpha = 0.15) +
  
  geom_ribbon(data = df_mid,
              aes(x = x, ymin = F_japan, ymax = 1),
              fill = "#4D6CFA", alpha = 0.3) +
  
  geom_ribbon(data = df_high,
              aes(x = x, ymin = F_japan, ymax = 1),
              fill = "#4D6CFA", alpha = 0.6) +
  # --- ABOVE RED (Japan) ---
  geom_ribbon(data = df_low,
              aes(x = x, ymin = F_japan, ymax = 1),
              fill = "#9E1946", alpha = 0.15) +
  
  geom_ribbon(data = df_mid,
              aes(x = x, ymin = F_japan, ymax = 1),
              fill = "#9E1946", alpha = 0.3) +
  
  geom_ribbon(data = df_high,
              aes(x = x, ymin = F_japan, ymax = 1),
              fill = "#9E1946", alpha = 0.6) +
  
  # --- BETWEEN CURVES ---
  geom_ribbon(data = df_low,
              aes(x = x,
                  ymin = pmin(F_japan, F_us),
                  ymax = pmax(F_japan, F_us)),
              fill = "#4D6CFA", alpha = 0.15) +
  
  geom_ribbon(data = df_mid,
              aes(x = x,
                  ymin = pmin(F_japan, F_us),
                  ymax = pmax(F_japan, F_us)),
              fill = "#4D6CFA", alpha = 0.3) +
  
  geom_ribbon(data = df_high,
              aes(x = x,
                  ymin = pmin(F_japan, F_us),
                  ymax = pmax(F_japan, F_us)),
              fill = "#4D6CFA", alpha = 0.6) +
  geom_line(aes(x=df_ecdf$x,y = df_ecdf$F_japan, color = "Japan: 30-44 years, male"), linewidth = 1.2) +
  geom_line(aes(x=df_ecdf$x,y = df_ecdf$F_us, color = "US: 30-44 years, male"), linewidth = 1.2) +
  scale_color_manual(name = "Country age-sex group",
                     values = c(
                       "Japan: 30-44 years, male" = "#9E1946",
                       "US: 30-44 years, male" = "#4D6CFA"
                     )) +
  
  labs(x = "Mass (kg)", y = "Cumulative population") +
  scale_y_continuous(limits=c(0,1.1), breaks=c(0,0.25,0.5,0.75,1), labels=c("0","2.5 mil","5 mil","7.5 mil", "1 mil"))+
  theme_minimal() + geom_vline(xintercept=25*1.7^2,color="grey30", linetype="dotted") +
  geom_vline(xintercept=30*1.7^2,color="grey30", linetype="dotted") + xlim(c(0,130))+
  theme(legend.position = "null") +
  geom_text(
    data = labels_df,
    aes(x = x, y = y, label = label),
    fontface = "bold",
    size = 5
  )+
  geom_segment(
    data = df_q,
    aes(x = 95, y = 1.035,
        xend = 90, yend = 0.995),
    color = "grey4",
    linetype="solid",
    arrow = arrow(length = unit(0.1, "cm"), type="closed")
  )


library(gridExtra)
library(grid)

table_df <- data.frame(
  Category = c("Optimal", "Current", "Excess"),
  `Total biomass` = c("A+B+C", "A+B+C+X+Y+Z", "X+Y+Z"),
  `Biomass due to overweight` = c("B+C", "B+C+Y+Z", "Y+Z"),
  `Biomass due to obesity` = c("C", "C+Z", "Z")
)

colnames(table_df) <- c(
  "",
  "Total biomass",
  "Biomass due to\noverweight",
  "Biomass due to\nobesity"
)

table_grob <-  tableGrob(
  table_df,
  rows = NULL,
  theme = ttheme_minimal(
    core = list(
      fg_params = list(col = "black"),
      bg_params = list(fill = NA, col = "black", lwd = 1)
    ),
    colhead = list(
      fg_params = list(fontface = "bold"),
      bg_params = list(fill = NA, col = "black", lwd = 1)
    )
  )
)
# Population 1 (row 2) → red
table_grob$grobs[table_grob$layout$t == 2] <- lapply(
  table_grob$grobs[table_grob$layout$t == 2],
  function(g) {
    g$gp$fill <- scales::alpha("#9E1946", 0.4)
    g
  }
)

# Excess (row 4) → blue
table_grob$grobs[table_grob$layout$t == 3] <- lapply(
  table_grob$grobs[table_grob$layout$t == 3],
  function(g) {
    g$gp$fill <- scales::alpha("#4D6CFA", 0.4)
    g
  }
)
first_col <- table_grob$layout$l == 1

table_grob$grobs[first_col] <- lapply(
  table_grob$grobs[first_col],
  function(g) {
    g$gp$font <- NULL
    g$gp$fontface <- "bold"
    g
  }
)
bottom_row <- max(table_grob$layout$t)
idx <- table_grob$layout$t == bottom_row &
  table_grob$layout$name == "core-fg"

table_grob$grobs[idx] <- lapply(
  table_grob$grobs[idx],
  function(g) {
    g$gp$font <- NULL          # avoid conflict with fontface
    g$gp$fontface <- "bold"
    g
  }
)
grid.newpage()
grid.draw(table_grob)




bmi_plot <- bmi_plot +
  theme(
    panel.border = element_rect(color = "grey40", fill = NA, linewidth = 0.5)
  )

height_plot <- height_plot +
  theme(
    panel.border = element_rect(color = "grey40", fill = NA, linewidth = 0.5)
  )

mass_plot <- mass_plot +
  theme(
    panel.border = element_rect(color = "grey40", fill = NA, linewidth = 0.5)
  )

cdf_plot <- cdf_plot +
  theme(
    panel.border = element_rect(color = "grey40", fill = NA, linewidth = 0.5)
  )

cdf_plot_areas <- cdf_plot_areas +
  theme(
    panel.border = element_rect(color = "grey40", fill = NA, linewidth = 0.5)
  )

library(patchwork)
left_col <- bmi_plot / mass_plot / cdf_plot
right_col <- cdf_plot_areas / wrap_elements(table_grob)
right_col <- cdf_plot_areas / wrap_elements(table_grob) +
  plot_layout(heights = c(2.5, 1))

final_plot <- left_col | plot_spacer() | right_col
final_plot +
  plot_layout(widths = c(1, 0.1, 1.6))

png("/Users/issylandray/Library/CloudStorage/OneDrive-Personal/Documents/UAH/Weightofnations/Code/Code_analysis2022/Output/Demonstrateshift.png",width=12,height=7, units = "in", res = 300)
final_plot +
  plot_layout(widths = c(1, 0.1, 1.6))
dev.off()
# 


df_ecdf$diff <- abs(df_ecdf$F_us - df_ecdf$F_japan)
df_ecdf$dx <- c(diff(df_ecdf$x), NA)
df_X <- df_ecdf[df_ecdf$x <= t25, ]
df_Y <- df_ecdf[df_ecdf$x > t25 & df_ecdf$x <= t30, ]
df_Z <- df_ecdf[df_ecdf$x > t30, ]
area_X <- sum(df_X$diff * df_X$dx, na.rm = TRUE)
area_Y <- sum(df_Y$diff * df_Y$dx, na.rm = TRUE)
area_Z <- sum(df_Z$diff * df_Z$dx, na.rm = TRUE)
1-area_X/(area_X+area_Y+area_Z)


df_ecdf$dx <- c(diff(df_ecdf$x), NA)

df_A <- df_ecdf[df_ecdf$x <= t25, ]
df_B <- df_ecdf[df_ecdf$x > t25 & df_ecdf$x <= t30, ]
df_C <- df_ecdf[df_ecdf$x > t30, ]

area_A <- sum((1 - df_A$F_japan) * df_A$dx, na.rm = TRUE)
area_B <- sum((1 - df_B$F_japan) * df_B$dx, na.rm = TRUE)
area_C <- sum((1 - df_C$F_japan) * df_C$dx, na.rm = TRUE)


area_A+area_B+area_C
area_B+area_C
area_C
area_A+area_B+area_C+area_X+area_Y+area_Z
area_B+area_C+area_Y+area_Z
area_C+area_Z
area_X+area_Y+area_Z
area_Y+area_Z
area_Z
```

## Supplementary Table 1: dataset with param estimates by age-sex-country group

```{r}
dta<-DF_ADULT_bayesmod_japan %>% full_join(DF_ADULT_bayesmod_USA)
head(dta)

dta$averagebodymass<- dta$`22biotot`/dta$TOTPOP
dta$averagebodymass_japan<- dta$`22biotot_JAPAN`/dta$TOTPOP
dta$averagebodymass_USA<- dta$`22biotot_USA`/dta$TOTPOP
table<-dta %>% dplyr::select(name, SEX, heightagegroups, TOTPOP, pivot, sigma1, sigma2, meanheight, sdheight, `22biotot`, averagebodymass, `22biotot_JAPAN`, averagebodymass_japan,`22biotot_USA`, averagebodymass_USA) %>% 
  dplyr::arrange(name, SEX, heightagegroups) %>% 
  dplyr::rename( Country = name, Sex = SEX, `Age group (years)` = heightagegroups, `Total Population (millions)` = TOTPOP, `BMI distribution pivot estimate` = pivot, `BMI distribution SD1 estimate (left-side SD)` = sigma1, `BMI distribution SD2 estimate (right-side SD)` = sigma2, `Mean Height` = meanheight, `Height SD` = sdheight, `Current total biomass (million kg)` = `22biotot`,`Current mean body mass (kg)` =averagebodymass, `Lowest scenario total biomass (million kg)`=`22biotot_JAPAN`, `Lowest scenario mean body mass (kg)`=averagebodymass_japan, `Highest scenario total biomass (million kg)`=`22biotot_USA`, `Highest scenario mean body mass (kg)`=averagebodymass_USA)

table$Sex <- if_else(table$Sex==0,"Women","Men")

177*2*6

saveRDS(table, "Output/Supplementary_Table1_resultsbycountryagesexgroup_17Jun.rds")
```
